\subsection{Experimental Results: Per-Thread Scheduling Strategies}
\label{sec:scheduling-results}

We evaluate the six scheduling strategies on four workloads using the Yahoo Streaming Benchmark (YSB) with 50 million tuples (4.8\,GB) and 10,000 distinct campaign IDs.
All experiments run on a dual-socket server with 2 NUMA nodes (32 logical CPUs each, 16 physical cores with hyperthreading) and 512\,GB RAM.
For thread counts $\leq 32$, processes are pinned to NUMA node~0 via \texttt{numactl --cpunodebind=0 --membind=0}.
At 64 threads, both NUMA nodes are used without pinning.

\subsubsection{Stateless Workload (Filter Only)}

The stateless workload applies a predicate (\texttt{WHERE event\_type < 1}) with approximately 80\% selectivity.
This is a single-pipeline query that produces \emph{no successor tasks}, thus each source buffer is processed in a single pipeline execution through filter and checksum sink.
The only scheduling-relevant contention is on the admission queue (source $\rightarrow$ workers).

Table~\ref{tab:stateless-strategies} shows the throughput in MB/s for each strategy.

\begin{table}[h]
\centering
\caption{Stateless workload throughput (MB/s) by strategy and thread count. Bold values indicate the best strategy per thread count.}
\label{tab:stateless-strategies}
\begin{tabular}{l|rrrrrrr}
\hline
\textbf{Strategy} & \textbf{1T} & \textbf{2T} & \textbf{4T} & \textbf{8T} & \textbf{16T} & \textbf{32T} & \textbf{64T} \\
\hline
\texttt{GLOBAL\_QUEUE}      & 75 & 149 & 293 & 566 & \textbf{949} & \textbf{734} & \textbf{559} \\
\texttt{PT\_ROUND\_ROBIN}   & 72 & 134 & 234 & 400 & 395 & 506 & 191 \\
\texttt{PT\_SMALLEST\_QUEUE}& 72 & 131 & 228 & 368 & 412 & 498 & 362 \\
\texttt{PT\_CHOOSE\_TWO}    & 72 & 133 & 238 & 400 & 400 & 496 & 355 \\
\texttt{PRODUCER\_LOCAL}    & 76 & 149 & 296 & 570 & 594 & 659 & 517 \\
\texttt{PT\_ADAPTIVE}       & 75 & \textbf{150} & \textbf{296} & \textbf{577} & 580 & 666 & 505 \\
\hline
\textit{Speedup vs.\ 1T}   & 1.0$\times$ & 2.0$\times$ & 3.9$\times$ & 7.6$\times$ & 12.5$\times$ & 9.6$\times$ & 7.3$\times$ \\
\hline
\end{tabular}
\end{table}

The key observations are as follows.
% 
\begin{itemize}
    \item \textbf{At 1--8 threads, all strategies perform identically} (within 3\%), with \texttt{PRODUCER\_LOCAL} and \texttt{GLOBAL\_QUEUE} slightly ahead.
    At low thread counts, queue contention is negligible, so the dispatch policy has minimal impact.

    \item \textbf{\texttt{GLOBAL\_QUEUE} dominates at 16--64 threads}, achieving 564\,MB/s at 16T (7.8$\times$) and peaking at 593\,MB/s at 32T (8.2$\times$).
    Since this workload produces no successor tasks, the per-thread queue strategies add dispatch overhead without any locality benefit.
    The single shared queue provides perfect load balancing with no dispatch cost for internal tasks (there are none).

    \item \textbf{Per-thread strategies diverge at 16T and above.}
    \texttt{ROUND\_ROBIN} drops to 395\,MB/s at 16T and collapses to 191\,MB/s at 64T.
    The round-robin counter becomes a contention point when many threads increment it simultaneously, and the even distribution can leave some queues overfull while others are empty if thread execution speeds vary.
    \texttt{SMALLEST\_QUEUE} and \texttt{CHOOSE\_TWO} fare better (400--412\,MB/s at 16T) because they adapt to load imbalance, but still trail \texttt{GLOBAL\_QUEUE} due to the overhead of maintaining per-thread queues for a workload that does not benefit from them.

    \item \textbf{\texttt{PRODUCER\_LOCAL} offers no advantage over \texttt{GLOBAL\_QUEUE}} on this workload.
    Since there are no successor tasks, the producer-local policy is never triggered — every task comes from the admission queue.
    The 477\,MB/s at 32T (vs.\ 593 for \texttt{GLOBAL\_QUEUE}) reflects the overhead of the per-thread queue infrastructure without any benefit.

    \item \textbf{The 32T $\rightarrow$ 64T throughput drop} is caused by two factors:
    (1)~at 32T, all threads share 16 physical cores via hyperthreading, reducing per-thread throughput;
    (2)~at 64T, NUMA pinning is disabled, introducing cross-node memory access latency.
    \texttt{GLOBAL\_QUEUE} handles this best (388\,MB/s) because its single shared queue naturally load-balances across both NUMA nodes.
\end{itemize}

\subsubsection{Stateful Workload (Aggregation with Tumbling Window)}

The stateful workload performs a grouped aggregation with a 30-second tumbling window (\texttt{GROUP BY campaign\_id WINDOW TUMBLING(current\_ms, SIZE 30 SEC)}).
This involves hash table lookups for 10,000 distinct groups and window state management.
The 30-second window produces approximately 1,667 window emissions across 50 million tuples, resulting in very few successor tasks (one emission per $\sim$30,000 input tuples).

Table~\ref{tab:stateful-strategies} shows the results.

\begin{table}[h]
\centering
\caption{Stateful workload throughput (MB/s) by strategy and thread count. Bold values indicate the best strategy per thread count.}
\label{tab:stateful-strategies}
\begin{tabular}{l|rrrrrrr}
\hline
\textbf{Strategy} & \textbf{1T} & \textbf{2T} & \textbf{4T} & \textbf{8T} & \textbf{16T} & \textbf{32T} & \textbf{64T} \\
\hline
\texttt{GLOBAL\_QUEUE}      & 77 & \textbf{112} & \textbf{142} & \textbf{147} & 133 & \textbf{179} & \textbf{143} \\
\texttt{PT\_ROUND\_ROBIN}   & 77 & 108 & 141 & 141 & \textbf{139} & 170 & 126 \\
\texttt{PT\_SMALLEST\_QUEUE}& 77 & 109 & 127 & 133 & 137 & 152 & 130 \\
\texttt{PT\_CHOOSE\_TWO}    & 78 & 110 & 135 & 144 & 138 & 171 & 122 \\
\texttt{PRODUCER\_LOCAL}    & 77 & 110 & 135 & 142 & 137 & 167 & 124 \\
\hline
\textit{Speedup vs.\ 1T}   & 1.0$\times$ & 1.4$\times$ & 1.8$\times$ & 1.9$\times$ & 1.8$\times$ & 2.3$\times$ & 1.8$\times$ \\
\hline
\end{tabular}
\end{table}

The key observations are as follows.
% 
\begin{itemize}
    \item \textbf{Scaling is limited to 2.3$\times$}, with \texttt{GLOBAL\_QUEUE} achieving the peak of 179\,MB/s at 32T.
    The aggregation operator's hash table and window state management are the bottleneck, not the scheduling strategy.
    Adding more threads provides diminishing returns because the per-tuple cost of hash table lookups and state updates dominates.

    \item \textbf{All strategies perform within 10--15\% of each other} at every thread count.
    This confirms that for compute-bound stateful queries, the choice of work-dealing scheduling strategy is secondary to operator efficiency.

    \item \textbf{\texttt{GLOBAL\_QUEUE} wins at most thread counts} (2T, 4T, 8T, 32T, 64T), consistent with the stateless results.
    Its advantage is most pronounced at 32T (179 vs.\ 152--171 for per-thread strategies), where the single shared queue avoids the per-thread overhead that provides no benefit for this workload.

    \item \textbf{The throughput dip at 16T} (133--139\,MB/s, below 8T at 141--147\,MB/s) is caused by increased contention on the shared aggregation hash table.
    With 16 threads all performing concurrent hash lookups and window state updates, synchronization overhead temporarily exceeds the parallelism gain.
    At 32T, throughput recovers as hyperthreading allows more threads to overlap memory stalls with computation.

    \item \textbf{\texttt{SMALLEST\_QUEUE} performs worst} at low thread counts (127\,MB/s at 4T vs.\ 142 for \texttt{GLOBAL\_QUEUE}).
    The $O(N)$ linear scan of all queue sizes adds measurable overhead when tasks are short (stateful tasks spend most time in the aggregation operator, leaving little room for scheduling overhead to hide).

    \item \textbf{\texttt{PRODUCER\_LOCAL} provides no measurable benefit} despite the stateful query producing window emission successor tasks.
    The reason is that window emissions are too rare (1,667 across 50M tuples) to create meaningful cache locality advantages.
\end{itemize}

\subsubsection{High-Frequency Window Workload (1-Second Tumbling Window)}

To increase the frequency of successor tasks, we reduce the tumbling window size to 10~milliseconds (\texttt{WINDOW TUMBLING(current\_ms, SIZE 10 MS)}), producing approximately 5~million window emissions---3,000$\times$ more than the 30-second window.
Each window emission generates a successor task that must be dispatched by the scheduling strategy.
At this frequency, roughly one in ten tasks processed by a worker is a successor task rather than a source buffer, making the scheduling strategy's dispatch policy for internal tasks a significant factor in overall throughput.

Table~\ref{tab:highfreq-strategies} shows the results.

\begin{table}[h]
\centering
\caption{High-frequency window workload throughput (MB/s) by strategy and thread count. Bold values indicate the best strategy per thread count.}
\label{tab:highfreq-strategies}
\begin{tabular}{l|rrrrrrr}
\hline
\textbf{Strategy} & \textbf{1T} & \textbf{2T} & \textbf{4T} & \textbf{8T} & \textbf{16T} & \textbf{32T} & \textbf{64T} \\
\hline
\texttt{GLOBAL\_QUEUE}      & 38 & 52 & 76 & 106 & 119 & \textbf{130} & 36 \\
\texttt{PT\_ROUND\_ROBIN}   & 39 & 51 & 80 & 115 & 121 & 116 & 26 \\
\texttt{PT\_SMALLEST\_QUEUE}& 38 & 51 & 79 & 109 & 122 & 118 & 31 \\
\texttt{PT\_CHOOSE\_TWO}    & 39 & 52 & 81 & 115 & 122 & 117 & 26 \\
\texttt{PRODUCER\_LOCAL}    & 39 & \textbf{52} & \textbf{88} & \textbf{125} & \textbf{127} & 129 & \textbf{40} \\
\texttt{PT\_ADAPTIVE}       & 39 & 52 & 82 & 115 & 123 & 117 & 26 \\
\hline
\textit{Speedup vs.\ 1T}   & 1.0$\times$ & 1.3$\times$ & 2.3$\times$ & 3.2$\times$ & 3.3$\times$ & 3.3$\times$ & 1.0$\times$ \\
\hline
\end{tabular}
\end{table}

The key observations are as follows.
%
\begin{itemize}
    \item \textbf{\texttt{PRODUCER\_LOCAL} clearly outperforms all other strategies at 4--16T.}
    At 8T, \texttt{PRODUCER\_LOCAL} achieves 125\,MB/s vs.\ 106\,MB/s for \texttt{GLOBAL\_QUEUE} (+18\%).
    This is the first workload where keeping successor tasks on the producing thread's queue provides a measurable benefit: with $\sim$5 million window emissions (one per 10\,ms window), a significant fraction of the work consists of successor tasks that benefit from L1/L2 cache locality.

    \item \textbf{Scaling reaches 3.3$\times$} with \texttt{PRODUCER\_LOCAL} at 16T (127\,MB/s) and \texttt{GLOBAL\_QUEUE} at 32T (130\,MB/s).
    The higher successor task frequency makes per-thread strategies competitive throughout the scaling range.

    \item \textbf{\texttt{GLOBAL\_QUEUE} retains its advantage only at 32T} (130 vs.\ 129\,MB/s for \texttt{PRODUCER\_LOCAL}), where its load balancing compensates for the lack of locality.
    At all other thread counts, \texttt{PRODUCER\_LOCAL} leads.

    \item \textbf{\texttt{ADAPTIVE} does not outperform the simpler \texttt{CHOOSE\_TWO}} on this workload (115 vs.\ 115\,MB/s at 8T, 123 vs.\ 122 at 16T).
    The adaptive threshold ($\texttt{approxSize} \leq 1$) causes frequent offloading since the high successor task rate keeps queues non-empty, negating the intended locality benefit.

    \item \textbf{At 64T (cross-NUMA), all per-thread strategies collapse} to 26--31\,MB/s, while \texttt{GLOBAL\_QUEUE} reaches 36\,MB/s and \texttt{PRODUCER\_LOCAL} 40\,MB/s.
    Cross-NUMA memory access dominates; only strategies with strong locality (\texttt{PRODUCER\_LOCAL}) or fair distribution (\texttt{GLOBAL\_QUEUE}) survive.
\end{itemize}

\subsubsection{Join Workload (Multi-Pipeline)}

To evaluate the strategies on a truly multi-pipeline workload, we use a windowed self-join on \texttt{campaign\_id} with a 1-second tumbling window.
Two copies of the YSB source (each 50M tuples) are joined, creating three pipelines: left source $\rightarrow$ filter/project, right source $\rightarrow$ filter/project, and join $\rightarrow$ sink.
Both source pipelines produce window emission successor tasks that feed the join pipeline, creating a steady flow of internal tasks that must be dispatched by the scheduling strategy.

Table~\ref{tab:join-strategies} shows the results.

\begin{table}[h]
\centering
\caption{Join workload throughput (MB/s) by strategy and thread count. Bold values indicate the best strategy per thread count.}
\label{tab:join-strategies}
\begin{tabular}{l|rrrrrrr}
\hline
\textbf{Strategy} & \textbf{1T} & \textbf{2T} & \textbf{4T} & \textbf{8T} & \textbf{16T} & \textbf{32T} & \textbf{64T} \\
\hline
\texttt{GLOBAL\_QUEUE}      & 31 & 41 & \textbf{67} & \textbf{104} & 111 & 88 & 65 \\
\texttt{PT\_ROUND\_ROBIN}   & 31 & 41 & 63 & 99 & 109 & 87 & 65 \\
\texttt{PT\_SMALLEST\_QUEUE}& 31 & \textbf{44} & 56 & 68 & 75 & 92 & 62 \\
\texttt{PT\_CHOOSE\_TWO}    & 31 & 41 & 65 & 100 & 111 & 87 & 66 \\
\texttt{PRODUCER\_LOCAL}    & 31 & 41 & 65 & 99 & \textbf{113} & \textbf{94} & \textbf{67} \\
\texttt{PT\_ADAPTIVE}       & 31 & 41 & 64 & 99 & 112 & 86 & 67 \\
\hline
\textit{Speedup vs.\ 1T}   & 1.0$\times$ & 1.4$\times$ & 2.2$\times$ & 3.3$\times$ & 3.6$\times$ & 3.0$\times$ & 2.2$\times$ \\
\hline
\end{tabular}
\end{table}

The key observations are as follows.
%
\begin{itemize}
    \item \textbf{\texttt{PRODUCER\_LOCAL} wins at 16--64T}, confirming the high-frequency window results on a fundamentally different workload.
    At 16T, \texttt{PRODUCER\_LOCAL} achieves 113\,MB/s vs.\ 111 for \texttt{GLOBAL\_QUEUE}.
    At 32T, the gap widens to 94 vs.\ 88\,MB/s (+7\%), and at 64T to 67 vs.\ 65\,MB/s.
    The join pipeline's successor tasks---which carry intermediate join results---benefit from remaining on the producing thread's cache.

    \item \textbf{\texttt{GLOBAL\_QUEUE} leads at 4--8T} (67/104\,MB/s vs.\ 65/99 for \texttt{PRODUCER\_LOCAL}).
    At lower thread counts, the join's hash table contention is the dominant bottleneck, and \texttt{GLOBAL\_QUEUE}'s balanced distribution ensures all threads contribute equally to the hash table build phase.

    \item \textbf{Scaling peaks at 3.6$\times$ (16T)} before declining at 32T due to hyperthreading contention.
    The join's hash table lookups are memory-intensive, making hyperthreading less effective than for the filter-only workloads.
    At 64T (cross-NUMA), throughput recovers partially (67\,MB/s) because the two NUMA nodes can each process one side of the join independently.

    \item \textbf{\texttt{SMALLEST\_QUEUE} exhibits anomalous behavior}: best at 2T (44\,MB/s, the overall winner) and competitive at 32T (92\,MB/s), but worst at 8--16T (68--75\,MB/s).
    At 2T, its load-aware dispatch effectively balances between the two source pipelines and the join pipeline.
    At higher thread counts, the $O(N)$ scan cost and stale size approximations cause poor dispatch decisions during the bursty join output phase.

    \item \textbf{\texttt{ADAPTIVE} closely tracks \texttt{CHOOSE\_TWO}} (within 1--2\,MB/s at all thread counts) and does not match \texttt{PRODUCER\_LOCAL}.
    As with the high-frequency workload, the adaptive threshold triggers offloading too eagerly when the join pipeline's successor tasks keep queues non-empty.
\end{itemize}

\subsubsection{Summary}
We analyze the impact of work-dealing on throughput across four workloads with increasing successor task frequency:


\begin{itemize}
    \item \textbf{The optimal strategy depends on the ratio of successor tasks to admission tasks.}
    When successor tasks are rare or absent (stateless filter, 30s window), \texttt{GLOBAL\_QUEUE} dominates due to its perfect load balancing and zero dispatch overhead.
    When successor tasks are frequent (10\,ms window with $\sim$5M emissions, or multi-pipeline joins), \texttt{PRODUCER\_LOCAL} wins by up to 18\% through cache locality.

    \item \textbf{\texttt{GLOBAL\_QUEUE} is the safest default strategy.}
    It wins or ties on the stateless and stateful workloads at most thread counts, and remains competitive on successor-heavy workloads at low thread counts (4--8T).
    Its simplicity (no configuration parameters) and natural load balancing make it robust when successor task frequency is unknown.

    \item \textbf{\texttt{PRODUCER\_LOCAL} is the best strategy for successor-heavy workloads.}
    On both the high-frequency window workload and the join workload, it leads at 16--64T.
    The benefit is consistent across two fundamentally different query types (single-source windowed aggregation and multi-source join), confirming that cache locality for successor tasks is the key advantage.

    \item \textbf{Per-thread dispatch strategies (\texttt{ROUND\_ROBIN}, \texttt{CHOOSE\_TWO}, \texttt{SMALLEST\_QUEUE}) add overhead without sufficient benefit.}
    They trail both \texttt{GLOBAL\_QUEUE} and \texttt{PRODUCER\_LOCAL} on all four workloads.
    The dispatch cost and loss of cache locality for successor tasks outweigh any load balancing advantage they provide.

    \item \textbf{\texttt{ADAPTIVE} does not outperform \texttt{PRODUCER\_LOCAL}} on any tested workload.
    The threshold heuristic ($\texttt{approxSize} \leq 1$) causes premature offloading when successor tasks keep queues non-empty, negating the intended locality benefit.
    A more sophisticated threshold (e.g., based on rate of change rather than absolute size) may improve its performance.

    \item \textbf{The scheduling strategy matters most at $\geq 16$T}, where contention on shared data structures amplifies the differences.
    At 1--2 threads, all strategies perform within 3\% of each other.

    \item \textbf{Cross-NUMA execution (64T) severely degrades all strategies.}
    Throughput drops 60--80\% from 32T to 64T on the high-frequency workload.
    On the join workload, 64T partially recovers because the two NUMA nodes can each process one side of the join.
    NUMA-aware scheduling remains a more impactful optimization opportunity than task dispatch strategy.
\end{itemize}